Model-order reduction of non-linear time-dependent neutronics via POD-Galerkin projection and DEIM

### 2022-06-07 paper_/main.tex
1st reviewer comment about steady state vs eigenvalue

	% ticks
	\foreach \x/\xtext in {15, 75, 105, 135, 165}
	\draw[black,xshift=0.1*\x cm] (0,.3) -- (0,0) node[below, font=\large] {$\xtext$};
	\foreach \y/\ytext in {15, 75, 105, 135, 165}
	\draw[black,yshift=0.1*\y cm] (.3,0) -- (0,0) node[left, font=\large] {$\ytext$};
	\end{scope}
	\end{tikzpicture}
	\caption{Schematic of LRA benchmark.}
	\label{fig:lra core}
\end{figure}
The transient is initiated by a control rod ejection with constant velocity over a period of 2 seconds leading to a super-critical, prompt transient, in which the power increases by about 10 orders of magnitude in a very short time.
The control rod movement is represented by changing the thermal absorption cross section of the control rod material.
In this application, the two-group diffusion equation represented in Sec.~\ref{sec:fom}, is coupled with thermal feedback through adiabatic fuel heat-up defined by
\begin{equation}
\frac{\partial}{\partial t}T(x,t) = \alpha[\Sigma_{f1}(x,t)\phi_1(x,t)+\Sigma_{f2}(x,t)\phi_2(x,t)] \, ,
\label{eq:heatup}
\end{equation}
and corresponding Doppler feedback defined by
\begin{equation}
\Sigma_{a1}(x,t) = \Sigma_{a1}(x,t=0)[1+\gamma(\sqrt{(T(x,t)}-\sqrt{(T_0)})] \, ,
\label{eq:doppler}
\end{equation}
where $\Sigma_{f_i}$ is the fission cross section of group $i$, $T$ is the temperature, $\Sigma_{a1}$ is the fast absorption cross section and $T_0$ is the initial temperature which is $300 $ K,
$\alpha = 3.83\times 10^{-11}~\text{K}~\text{cm}^3$ and $\gamma = 3.034\times 10^{-3} ~\text{K}^{-0.5}$.
The power local is computed as
\begin{equation}
P(x,t)=\kappa [\Sigma_{f1}(x,t)\phi_1(x,t)+\Sigma_{f2}(x,t)\phi_2(x,t)] \, ,
\end{equation}
where $\kappa = 3.204\times 10^{-11}~ \text{W/fission}$.
The model was implemented in the deterministic transport code Detran~\cite{roberts2014advanced}.
A mesh-centered, finite-difference discretization was used in space, while a first-order, backward-difference temporal discretization was used with a fixed 0.001 s step size.  
The initial condition for the transient was computed by solving the steady-state equation, for which $k_{\text{eff}}$ was found to be 0.9975.
Note that at each time step the fast absorption cross section is a function of the solution itself.
The convergence criterion used to reduce the $\text{L}_2$ norm  of the relative flux error between two consecutive iterations to less than $1\times 10^{-6}$, or
\begin{equation}
\|\frac{\mathbf{\Phi}^{i+1} - \mathbf{\Phi}^{i}} {\mathbf{\Phi}^i}\| \le  1\times 10^{-6} \, ,
\label{eq:convergance criteria} 
\end{equation}
where the superscript $i$ denotes the iteration number.
The transient was calculated for 3.0 seconds.
Figure~\ref{fig:lra fom power} shows the reference full-order model power along with the core average temperature.
This reference solution is not necessarily numerically converged in either space or time but represents the simulation response that one wishes to approximate.
\begin{figure}[h!]
	\includegraphics[width=1.0\linewidth]{../figures/LRA_fom_power_temperature.pdf}
	\caption{LRA core power}
	\label{fig:lra fom power}
\end{figure}

@@

\begin{figure}[H]
	\includegraphics[width=1.0\linewidth]{../figures/_.pdf}
	\caption{Relative error of the power predicted by MDEIM-ROM}
	\label{fig:deim power error}
\end{figure}
To understand the impact of employing the MDEIM on the error, the results are compared with these of a direct-projection ROM, i.e., with performing direct operator projection at each time step.
The relative difference of the power between both models is shown in Fig.~\ref{fig:rom-deim power error}.
The shown errors implies that MDEIM accurately approximates the problem operator.

>>>>

	% ticks
	\foreach \x/\xtext in {15, 75, 105, 135, 165}
	\draw[black,xshift=0.1*\x cm] (0,.3) -- (0,0) node[below, font=\large] {$\xtext$};
	\foreach \y/\ytext in {15, 75, 105, 135, 165}
	\draw[black,yshift=0.1*\y cm] (.3,0) -- (0,0) node[left, font=\large] {$\ytext$};
	\end{scope}
	\end{tikzpicture}
	\caption{Schematic of LRA benchmark.}
	\label{fig:lra core}
\end{figure}
The transient is initiated by a control rod ejection with constant velocity over a period of 2 seconds leading to a super-critical, prompt transient, in which the power increases by about 10 orders of magnitude in a very short time.
The control rod movement is represented by changing the thermal absorption cross section of the control rod material.
In this application, the two-group diffusion equation represented in Sec.~\ref{sec:fom}, is coupled with thermal feedback through adiabatic fuel heat-up defined by
\begin{equation}
\frac{\partial}{\partial t}T(x,t) = \alpha[\Sigma_{f1}(x,t)\phi_1(x,t)+\Sigma_{f2}(x,t)\phi_2(x,t)] \, ,
\label{eq:heatup}
\end{equation}
and corresponding Doppler feedback defined by
\begin{equation}
\Sigma_{a1}(x,t) = \Sigma_{a1}(x,t=0)[1+\gamma(\sqrt{(T(x,t)}-\sqrt{(T_0)})] \, ,
\label{eq:doppler}
\end{equation}
where $\Sigma_{f_i}$ is the fission cross section of group $i$, $T$ is the temperature, $\Sigma_{a1}$ is the fast absorption cross section and $T_0$ is the initial temperature which is $300 $ K,
$\alpha = 3.83\times 10^{-11}~\text{K}~\text{cm}^3$ and $\gamma = 3.034\times 10^{-3} ~\text{K}^{-0.5}$.
The power local is computed as
\begin{equation}
P(x,t)=\kappa [\Sigma_{f1}(x,t)\phi_1(x,t)+\Sigma_{f2}(x,t)\phi_2(x,t)] \, ,
\end{equation}
where $\kappa = 3.204\times 10^{-11}~ \text{W/fission}$.
The model was implemented in the deterministic transport code Detran~\cite{roberts2014advanced}.
A mesh-centered, finite-difference discretization was used in space, while a first-order, backward-difference temporal discretization was used with a fixed 0.001 s step size.  
The initial condition for the transient was computed by solving the eigenvalue problem, for which $k_{\text{eff}}$ was found to be 0.9975.
Note that at each time step the fast absorption cross section is a function of the solution itself.
The convergence criterion used to reduce the $\text{L}_2$ norm  of the relative flux error between two consecutive iterations to less than $1\times 10^{-6}$, or
\begin{equation}
\|\frac{\mathbf{\Phi}^{i+1} - \mathbf{\Phi}^{i}} {\mathbf{\Phi}^i}\| \le  1\times 10^{-6} \, ,
\label{eq:convergance criteria} 
\end{equation}
where the superscript $i$ denotes the iteration number.
The transient was calculated for 3.0 seconds.
Figure~\ref{fig:lra fom power} shows the reference full-order model power along with the core average temperature.
This reference solution is not necessarily numerically converged in either space or time but represents the simulation response that one wishes to approximate.
\begin{figure}[h!]
	\includegraphics[width=1.0\linewidth]{../figures/LRA_fom_power_temperature.pdf}
	\caption{LRA core power}
	\label{fig:lra fom power}
\end{figure}

@@

%\begin{figure}[H]
%	\includegraphics[width=1.0\linewidth]{../figures/_.pdf}
%	\caption{Relative error of the power predicted by MDEIM-ROM}
%	\label{fig:deim power error}
%\end{figure}
To understand the impact of employing the MDEIM on the error, the results are compared with these of a direct-projection ROM, i.e., with performing direct operator projection at each time step.
The relative difference of the power between both models is shown in Fig.~\ref{fig:rom-deim power error}.
The shown errors implies that MDEIM accurately approximates the problem operator.
